\blue{...Numerical investigation of viscous flow instabilities through heterogeneous porous media using adaptive and unstructured meshes under different mobility scenarios...}

### 2017-07-09 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
do not rush kostas

%%%%%%    
%\red{ NEEDS TO BE MOVED LATER In Figures \ref{fem_cv_represent_vr1_vr10} there is a different set of simulations. There is a comparison between the $VR=1$ and $VR=10$ case and the reason was to imply the same domain, boundary and initial conditions and number of elements to underastand how the cross-flow will develop. The purpose of these two cases was to verify that the model is working for cross flow cases thus will allow to proceed further and expand this technique into more complex formations.}    
%%%%%%

@@

Figures~\ref{fig:HeleShawHeter_VR10} and ~\ref{fig:HeleShawHeter_VR150} show formation of multi-scale elongations with continuous growth and coalescence of dendritic finger branching in simulations performed with viscosity ratio of 10 and 150, respectively. In both simulations, maximum fingers width of approximately 0.44-0.56 ({\it VR}=10) and 0.24-0.44 cm were found, whereas estimated values (based on Eqn.~\ref{eqn:guaneqn}) were 0.25-4.50 ({\it VR}=10) and 0.88-3.00 cm (Fig.~\ref{fig:homoheleshaw_VN10_VN150} c and d, Table~\ref{table:Heleshaw}). Such calculated ranges \citep[\ie theoretical values based on analytic solutions due to][]{mclean_1981} clearly overestimate fingers dimensions as the mathematical formulation (and therefore Eqn.~\ref{eqn:guaneqn}) assumes incompressible flows in homogeneous porous medium, \ie it does not take into account any spatial variability of heterogeneity.

@@

Mesh adaptivity methods have been extensively used by the computational geofluid dynamic community to help capturing detailed flow dynamics, compositional non-equilibrium fluid displacement and solid-fluid interactions \citep{pluszny_2007,pietro_2014,su_2016,melnikova_2016}. In such methods, the mesh grid is continuously modified (\ie adjustments of the number and distribution of the degrees of freedom to reduce solution error) to focus resolution where is necessary as the simulation evolves in time and space. There are four main families of mesh adaptivity methods: adaptive mesh refinement (AMR), edge and face element manipulations whereas keeping the order of the element basis function fixed (h-adaptivity), mesh deformation (r-adaptivity) and changes of element basis function order (p-adaptivity). Detailed description and analysis of these mesh adaptivity methods are beyond the scope of this manuscript but can be readily found in \citet{lo_book} \citep[see also][]{plewa_book,frey_book}.

The mesh adaptive algorithm embedded in the Fluidity/IC-FERST model utilises a metric tensor field dependent on solution interpolation error-estimates which locally control the topology of elements in the metric \citep{pain_2001,power_2006}. Mesh optimisation generates unstructured finer mesh in regions where flow properties change faster and coarser mesh in regions where properties change more slowly \citep{piggott_2006}.

Having already verified that model is working for the hele-shaw experimental test-cases, now a more complicated case will be introduced with regions of different geometrical domains and different permeabilities. Initial and boundary conditions are described in fig.\ref{fig:testcase_heter_domain}, while inlet velocity is $u=1$ on the left hand side and the pressure set to $0$ on the outlet. A quick reminder at this point is that the simulator we are testing has also options to introduce gravity and/or corriolis, but for this current set of simulations these are neglected. What is expected is a significant growth of fingers when comparing mobility ratios that differ an order of magnitude. %(fig.\ref{fig:1testcase})
The next step is a comparison under the same mobility ratio of $10$, using the fixed and the adaptive mesh. For the fixed mesh we have $4600$ elements, but for the adaptive mesh we start with $400$ up to $600$, depends how fine we would like the mesh to be, and we reach up to $24000$. %Figure \ref{fig:testcase_heter_domain} gives a description of the domain that will be used.  
Figures \ref{fig:2testcase_a}%\ref{fig:2testcase_b}, \ref{fig:2testcase_c} 
-\ref{fig:2testcase_d} will give a description of the simulation for different time-steps as these are described. In this set the interfacial instability is becoming more profound as the flow develops. Based on these two cases so far, an initial relation between the quantitative and qualitative outcome of our simulations can be derived. Fingers will develop faster in areas where permeabilities are higher. This can be seen from the overall sharp increase of velocity magnitude during the injection of the fluid $1$.

>>>>

@@

\medskip
Figures~\ref{fig:HeleShawHeter_VR10} and ~\ref{fig:HeleShawHeter_VR150} show formation of multi-scale elongations with continuous growth and coalescence of dendritic finger branching in simulations performed with viscosity ratio of 10 and 150, respectively. In both simulations, maximum fingers width of approximately 0.44-0.56 ({\it VR}=10) and 0.24-0.44 cm were found, whereas estimated values (based on Eqn.~\ref{eqn:guaneqn}) were 0.25-4.50 ({\it VR}=10) and 0.88-3.00 cm (Fig.~\ref{fig:homoheleshaw_VN10_VN150} c and d, Table~\ref{table:Heleshaw}). Such calculated ranges \citep[\ie theoretical values based on analytic solutions due to][]{mclean_1981} clearly overestimate fingers dimensions as the mathematical formulation (and therefore Eqn.~\ref{eqn:guaneqn}) assumes incompressible flows in homogeneous porous medium, \ie it does not take into account any spatial variability of heterogeneity.

@@

\medskip
Mesh adaptivity methods have been extensively used by the computational geofluid dynamic community to help capturing detailed flow dynamics, compositional non-equilibrium fluid displacement and solid-fluid interactions \citep{pluszny_2007,pietro_2014,su_2016,melnikova_2016}. In such methods, the mesh grid is continuously modified (\ie adjustments of the number and distribution of the degrees of freedom to reduce solution error) to focus resolution where is necessary as the simulation evolves in time and space. There are four main families of mesh adaptivity methods: adaptive mesh refinement (AMR), edge and face element manipulations whereas keeping the order of the element basis function fixed (h-adaptivity), mesh deformation (r-adaptivity) and changes of element basis function order (p-adaptivity). Detailed description and analysis of these mesh adaptivity methods are beyond the scope of this manuscript but can be readily found in \citet{lo_book} \citep[see also][]{plewa_book,frey_book}.

\medskip
The mesh adaptive algorithm embedded in the Fluidity/IC-FERST model utilises a metric tensor field dependent on solution interpolation error-estimates which locally control the topology of elements in the metric \citep{pain_2001,power_2006}. Mesh optimisation generates unstructured finer mesh in regions where flow properties change faster and coarser mesh in regions where properties change more slowly \citep{piggott_2006}.

\medskip
Having already verified that model is working for the hele-shaw experimental test-cases, now a more complicated case will be introduced with regions of different geometrical domains and different permeabilities. Initial and boundary conditions are described in fig.\ref{fig:testcase_heter_domain}, while inlet velocity is $1$ on the left hand side, and the pressure is set to $0$ on the outlet. What is expected is a significant growth of fingers when comparing mobility ratios that differ an order of magnitude. 
%(fig.\ref{fig:1testcase})
Figure \ref{fem_cv_represent_vr1_vr10} describes an initial comparison between the $VR=1$ and $VR=10$ case under the same domain, boundary and initial conditions and number of elements so as to underastand how the cross-flow in the upper layers can be developed compared with the fingers that will appear below. The purpose of these two cases was to verify that the model is working for cross flow cases thus will allow to proceed further and expand this technique into more complex formations. The next step is a comparison under the same mobility ratio of $10$, using the fixed and the adaptive mesh. For the fixed mesh we have $4600$ elements, but for the adaptive mesh we start with $400$ up to $600$, depends how fine we would like the mesh to be, and we reach up to $24000$. %Figure \ref{fig:testcase_heter_domain} gives a description of the domain that will be used.

\medskip
Figures \ref{fig:2testcase_a}%\ref{fig:2testcase_b}, \ref{fig:2testcase_c} 
-\ref{fig:2testcase_d} give a summary of the simulation for different time-steps as these are described. In this set, the interfacial instability is becoming more profound as the flow develops. Based on these two cases so far, an initial relation between the quantitative and qualitative outcome of our simulations can be derived. Fingers will develop faster in areas where permeabilities are higher. This can be seen from the overall sharp increase of velocity magnitude during the injection of the fluid $1$.
